Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-08 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:08	18:24:38	(Tue 8.Sep)

\begin{figure*}[t]\centering
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel.  Note how LCA prefers a rectangular window ($\beta=0$) on the anchor.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_beta.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. ]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_phi.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{Kaiser windows chosen for each pixel in the image. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
\label{fig_sim}
\end{figure*}

@@

The ultimate goal of adaptive beamformers is to take the information available in the wavefield and use it to either improve image resolution, noise suppression, or both. This non-linear processing increases the image bandwidth introduces a need for a sampling rate higher than the Nyquist rate. For MVDR a factor 10 oversampling laterally compared to the Nyquist rate is often needed to ensure  spatialhift-invariance in source amplitude~\cite{Asen2014}. Since LCA operates with the same optimization criterion, we will be using approximately a factor 8 lateral oversampling for both beamformers.

\todo{Need $M$ in the image sector. Why???}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2009}. It is a high resolution synthetic aperture sonar with array length 1.2\;m, 100\;kHz operating frequency, a bandwidth of 40\;kHz and 25$^\circ$ opening angle.

@@

In the image reconstruction we have run MVDR with a fairly aggressive yet stable set of parameters; subarray size $L=16$ and $d=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging.

@@

Observe from \Fig{mvdr_selected_windows} that MVDR seems prefers symmetric window responses, even if it is free to choose non-symmetric ones. The symmetry is most predominant in shadow and speckle regions, while in highlight the windows are approximately symmetric within the illuminated\todo{valid term?} sector of the seafloor. We have observed this symmetry to be a side-effect of running MVDR in near snapshot mode in an active system, due to the steps needed to build the sample covariance matrix.If we look at the peak sidelobe levels of the MVDR windows they are for the most part between -13\,dB and -43\,dB. This corresponds to that of the rectangular and Hamming window, or a Kaiser window with $\beta=0$ and $\beta=2$, respectively.

We hypothesize that LCA should work even if it only selects from Kaiser windows. If they span a decent range of $\beta$-values, and some steering variations is applied to each, we should have responses that resembles those in \Fig{mvdr_selected_windows}. The Kaiser window is easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian. It is also optimal in the sense of having its peak power concentration near the steering angle.

@@

What is the best set of Kaiser windows to use with LCA, and how much should they be steered? To investigate this we constructed large and equally sized window databases containing windows with varied upper boundaries for $\beta$ and $\phi$. As lower boundaries we chose $\beta=0$ and $\phi=\pm0^\circ$, which includes the rectangularly weighted DAS image. The images produced are shown in \Fig{oversampling_mosaic_bounds}.

We suggest an upper bound for $\beta$ between 2 and 5, where 5 is slightly better at suppressing sidelobes. For $\phi$ we suggest to steer either to 50\% or 100\% of the -3dB width, which leads to a resolution of 73\% or 54\% of the rectangular window, respectively. As lower boundaries we choose $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.

>>>>

\begin{figure*}[t]\centering
\subfloat[$\beta$-values of the Kaiser window chosen for each image pixel. Note how LCA prefers a a narrow response ($\beta=0$) on the anchor, and on other strong sources with little lateral interference. On the sources where lateral interference is present it chooses the widest response with the best sidelobe suppression.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_beta.pdf}%%
\label{oversampling_mosaic_bounds_beta}}
\hfil
\subfloat[$\phi$-values of the Kaiser window chosen for each image pixel. On strong sources LCA prefers selects windows that are steered away from the source. This is what improves the FWHM measurement in \Fig{oversampling_mosaic_bounds}.]{\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds_lca_windows_phi.pdf}%
\label{oversampling_mosaic_bounds_phi}}
\caption{Kaiser windows chosen for each pixel in the image. The underlying image is the one shown in \Fig{oversampling_mosaic_bounds}, with the same parameters. The location of the anchor, cut line and main scatter locations of the cross is marked in red.}
\label{fig_sim}
\end{figure*}

@@

The ultimate goal of adaptive beamformers is to take the information available in the wavefield and use it to either improve image resolution, noise suppression, or both. This non-linear processing increases the image bandwidth introduces a need for a sampling rate higher than the Nyquist rate. For MVDR a factor 10 oversampling laterally compared to the Nyquist rate is often needed to ensure  spatial shift-invariance in source amplitude~\cite{Asen2014}. Since LCA operates with the same optimization criterion, we will be using approximately a factor 8 lateral oversampling for both beamformers.

% \todo{Need $M$ in the image sector. Why???}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2009}. It is a high resolution synthetic aperture sonar with 1.2\;m array length, 100\;kHz operating frequency, 40\;kHz bandwidth and 25$^\circ$ opening angle on transmission.

@@

In the image reconstruction we have run MVDR with subarray length $L=16$ and $d=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging. This is a fairly aggressive yet stable set of parameters.

@@

Observe from \Fig{mvdr_selected_windows} that MVDR seems prefers symmetric window responses, even if it is free to choose non-symmetric ones. The symmetry is most predominant in shadow and speckle regions, while in highlight the windows are approximately symmetric within the illuminated\todo{valid term?} sector of the seafloor. We have observed this symmetry to be a side-effect of the averaging steps needed to build the sample covariance matrix, which is key to allow  MVDR to run in near snapshot mode in an active system. The peak sidelobe levels of the MVDR windows they are mostly between -13\,dB and -43\,dB. This corresponds to that of the rectangular and Hamming window, or a Kaiser window with $\beta=0$ and $\beta=2$, respectively.

From this we hypothesize that LCA should work even if it only selects from various Kaiser windows. If they span a decent range of $\beta$-values, and some steering variations is applied to each, we should have responses that resembles those in \Fig{mvdr_selected_windows}. The Kaiser window is easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian. It is also optimal in the sense of having its peak power concentration near the steering angle.

@@

What is the best set of Kaiser windows to use with LCA, and how much should they be steered? To investigate this we constructed large and equally sized window databases containing windows with varied upper boundaries for $\beta$ and $\phi$. As lower boundaries we chose $\beta=0$ and $\phi=0^\circ$, which includes the rectangularly weighted DAS image. The images produced are shown in \Fig{oversampling_mosaic_bounds}. The lateral distance $\Delta x$ between the -3dB points of the anchor is measured for each image, with the reference $\Delta x_\text{ref}$ being the rectangularly weighted DAS image. This measure is also called the full width half maximum (FWHM). In the case of a imaging a point scatter, it is closely related to the resolution of of the system. For a rectangular window, for instance, the resolution is 0.89$\cdot$FWHM.

The images show much improved sidelobe suppression with a an upper for bound for $\beta$ of 2, but it improved further as this bound was adjusted up to 5. After 5 no noticeable difference was observed. The images also demonstrate that increasing the steering $\phi$ improves the resolution of the strong scatterers. For this sonar the Rayleigh resolution 0.72$^\circ$, or roughly 24\,cm at the imaging range of 19\,m. we suggest to steer either to 50\% or 100\% of the -3dB width, which leads to a resolution of 73\% or 54\% of the rectangular window, respectively. As lower boundaries we choose $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.
